---
title: "Part I"
output:
  word_document: default
  pdf_document: default
---

## Excercise 5 (B)

# Absract
In this excersise we analyzed the data-set "ask1mf.xls" which refer to a sample of 32 observations of high school students. The variables in this data-set are GPA (mean grade), TUCE (exams grade in finance), PSI (student relation with finance). We were called to perform a scatter plot between the variables, find the sample mean and sample variance, find the correlation coefficient, implement linear regression and calculate the anova matrix.

# Implementation
At start, we load the required libraries. Then, we read the data-set with the read_excel function which puts the data into a data-frame. Then, we view the column names and the first few rows to get an idea on how the data-set looks like. The summary function gives us some basic descriptives of the data (min value, max value, median, mean, 1st Qu, 3rd Qu). After that, we do a scatter plot of the variable GPA vs variable TUCE.

```{r}
# load libraries
library(tidyverse)
library(ggpubr)
library(aod)
library(ggplot2)
library(readtext)
library(rgl)
library(dplyr)
library(readxl)

# read data
mydata <- read_excel("ask1mf.XLS")

# view what is on the data frame
names(mydata)

# view the first few rows of the data
head(mydata)

# view basic descriptives of the data
summary(mydata)

# scatter plot of GPA vs TUCE
plot(mydata$GPA, mydata$TUCE, main="Scatter plot of GPA vs TUCE",
     xlab="GPA ", ylab="TUCE", pch=19)
```

From the scatter plot output above we don't see a pattern/relationship between the variables GPA and TUCE. There is no ascending or descending relationship between the variables as we go from right to left on the plot.

As a next step, make two groups. Group 1 involves the variables GPA+PSI and Group 2 involves the variables TUCE+PSI. Then we implement a scatter plot between Group 1 (GPA+PSI) and Group 2 (TUCE+PSI).

```{r}
# scatter plot of Group 1 (GPA + PSI) vs Group 2 (TUCE + PSI)
plot(mydata$GPA + mydata$PSI, mydata$TUCE + mydata$PSI, main="Scatter plot of Group 1 (GPA + PSI) vs Group 2 (TUCE + PSI)",
     xlab="Group 1 (GPA + PSI)", ylab="Group 2 (TUCE + PSI)", pch=19)
```

Now the scatter plot presents an ascending relationship between the two groups. As we go from left to right the two groups increase together.

Furthermore, we calculate the sample mean and sample variance for each variable. Function summary shows us the sample mean for each variable and function var presents the sample variance for each variable.

```{r}
# view sample mean
summary(mydata)
# view sample variance
var_GPA <- var(mydata$GPA)
var_GPA
var_TUCE <- var(mydata$TUCE)
var_TUCE
var_PSI <- var(mydata$PSI)
var_PSI
```

Additionally, we calculate the correlation coefficient between variables GPA and TUCE.

```{r}
# correlation coefficient between variables GPA and TUCE
cor(mydata$GPA, mydata$TUCE, method = "kendall")
```

In this step we calculate the linear regression between the variables TUCE AND GPA. TUCE is the dependent variable whereas GPA is the independent one.

```{r}
# linear regression
# GPA is the independent variable whereas TUCE is the dependent variable
lmmodel <- lm(mydata$TUCE ~ mydata$GPA, data = mydata)
print(lmmodel)

# view the statistical summary of the linear regression model
summary(lmmodel)

# correlation coefficient between GPA and TUCE
cor(mydata$GPA, mydata$TUCE)

# scatter plot of the linear regression line
ggplot(mydata, aes(GPA, TUCE)) + geom_point() + stat_smooth(method = lm) + ggtitle("GPA vs TUCE (Linear regression model)")

```

The linear model coefficients intercept is w0 = 11.853 and slope is w1 = 3.235. The multiple R-squared is 14.98%.
We have a low R-squared value which means that we have a high bias on our data-set. The linear model scatter plot verifies the scatter plot assumptions we did in earlier step between variables TUCE and GPA. There is no obvious relationship between these two variables. As a last step, we calculate the anova matrix between the variables TUCE and GPA.
```{r}
# anova matrix
fit <- aov(TUCE ~ GPA, data=mydata) 
fit

```

