---
title: "Part II"
output:
  word_document: default
  pdf_document: default
---

## Excercise 7.2

# Absract
In this exercise we analyzed a data-set "insurance.csv" which includes 1338 samples on variables: age, sex, bmi, children, smoker, region and charges. We are called to investigate how these variables affect the charges variable.


# Implementation
First we read the data with read.csv function and put it in a data-frame. Then, we view the column names and the first few rows to get an idea on how the data-set looks like. The summary function gives us some basic descriptives of the data (min value, max value, median, mean, 1st Qu, 3rd Qu).

```{r}
# load libraries
library(tidyverse)
library(ggpubr)
library(aod)
library(ggplot2)
library(readtext)
library(rgl)
library(gplots)
library(corrplot)
library(readxl)
library(psych)
library(interplot)

# read data
mydata <- read.csv("insurance.csv", header=TRUE)

# view what is on the data frame
names(mydata)

# view the first few rows of the data
head(mydata)

# view basic descriptives of the data
summary(mydata)
```

Then we make a check of the variables with the "psych" package. We extract the geometric mean and also the harmonic mean for each of the variables.

```{r}
# geometric.mean function
geometric.mean(mydata[c(1,3,4,7)],na.rm=TRUE)

# harmonic.mean function
harmonic.mean(mydata[c(1,3,4,7)],na.rm=TRUE)
```

After that, we perform the t.test function to check how each numeric variable affects the charges variable.

```{r}
age <- mydata$age
bmi <- mydata$bmi
children <- mydata$children
charges <- mydata$charges

# two sample t-test to check how each variable affects charges
t.test(age,charges)
t.test(bmi,charges)
t.test(children,charges)
```

In every t.test analysis the p-value is 0 which means that all variables are statistically significantly associated with each other. Therefore, all variables affect the charges variable.
